---
title: 'Worksheet: Introduction to Plotting with ggplot2'
author: "Damien Ronk"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output:
  html_document:
    theme: yeti
    highlight: textmate
    toc: yes
    toc_depth: 4
  pdf_document:
    toc: yes
    toc_depth: '4'
subtitle: STT 2860 Intro to Data Mgmt & Visualization
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

***

### Packages Used

```{r}
# load all packages used here
library(tidyverse)
library(nycflights13)
```


### Datasets Used

#### Popular Kids

The dataset `popular.csv` is from [Dr. Thomley's Data Files](https://stat-jet-asu.github.io/Datasets/DatasetList.html). You can read about it [HERE](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/popular.html).

```{r}
# import the dataset from the web URL
kids <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/popular.csv")
```

```{r}
# explore the dataset
# can do so in the environment
# or we can explicitly ask for information
glimpse(kids)
head(kids)
```

#### Semmelweis Data

The dataset `semmelweisannualbyclinic.csv` is from [Dr. Thomley's Data Files](https://stat-jet-asu.github.io/Datasets/DatasetList.html). You can read about Dr. Ignaz Semmelweis [HERE](https://www.pbs.org/newshour/health/ignaz-semmelweis-doctor-prescribed-hand-washing). This dataset contains annual observations for two different clinics that led him to implement his handwashing experiment.

```{r}
semm <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/semmelweisannualbyclinic.csv")
```

```{r}
# explore the dataset
glimpse(semm)
head(semm)
```

#### Flights Data

The `flights` dataset is from the nycflights13 package, which is installed on the RStudio server.

```{r}
# save the dataset to your Environment
flights <- flights
```

```{r}
# explore the dataset
glimpse(flights)
head(flights)
```


### Using `geom_bar()`

#### Basic Bar Plot

```{r}
ggplot(kids)
# bare plot
```

```{r}
ggplot(kids, aes(x = top_goal))
# we've told it what we want to plot, and that its associated with the x axis
```

```{r}
# plus sign must be at the end of this line, can not be on second line
ggplot(kids, aes(x = top_goal)) +
  geom_bar() #must put parenthesis because it is a FUNCTION
# now we've told it what kind of plot we actually want
```

```{r}
ggplot(kids, aes(x = district_type)) +
  geom_bar()
```

```{r}
ggplot(kids, aes(x = gender)) +
  geom_bar()
```

Now we want our code to be more informative! 

#### Fancy Bar Plots 

```{r}
# horizontal bar plot
ggplot(kids, aes(y = top_goal)) +
  geom_bar()
#same as the bare bones, but change x to y
```

```{r}
# use a different layer using `coord_flip`  
ggplot(kids, aes(x = top_goal)) +
  geom_bar() +
  coord_flip()

# does the same thing as changing the x and y value, this is more common
```

```{r}
# adding useful labels!
ggplot(kids, aes(x = top_goal)) +
  geom_bar() +
  labs(
    title = "Top Goal in School Picked by Sampled Students",
    subtitle = "* participants chose one goal from three options",
    x = "Top Goal Options",
    y= "Number of Students",
    caption = "data file: popular.csv"
  )
# plot with informative labels! later, we will use theme() to make it look nicer
```

### Using `geom_col()`

#### Subheader Here!



***
```{r}
sessionInfo()
```
